Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-04 main.tex
Updates from Overleaf

\usepackage{hyperref}
\usepackage{xr}
% for xr in your preamble
% https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
% my external document I would like to reference the labels of. 
\myexternaldocument{Supplementary_figures} %.tex .aux

@@

%\section{abstract}
%%%%%%%%%%
% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{The survival analysis of the Cancer Genome Atlas (TCGA) dataset is a well-known method to discover the gene expression-based prognostic biomarkers of head and neck squamous cell carcinoma (HNSCC). It is necessary to determine a cutoff point by the patients' dichotomization for the continuous gene expression in survival analysis. There is some optimization software for cutoff determination. However, the software's predetermined cutoffs are usually set at the median or quantiles of RNA sequencing value to perform the survival analyses. There are few clinicopathological features available on their pre-processed data sets.
We developed a workflow that includes data retrieving and pre-processing, feature selection, sliding-window cutoff mining, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
Using this workflow on the TCGA HNSCC cohort, we scanned human protein-coding genes. After adjustment with confounders, we found that the clinical tumor stage and the surgical margin involvement are independent risk factors in patient survival. According to the resulting tables with Bonferroni adjusted P-value under the optimal cutoff and the hazard ratio, three biomarker candidates, CAMK2N1, CALML5, and FCGBP, are significantly associated with the patients' overall survival. The other independent HNSCC data were also used to validate this discovery.
We suggest the transcriptomic analysis could help for biomarker discovery.}

@@

They employed the Database for Annotation, Visualization, and Integrated Discovery (DAVID, available at https://david.ncifcrf.gov/) to obtain information for Gene Ontology (GO), including biological processes, the cellular component, and molecular function. 
The Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway analysis was also used to annotate the potential functions.

@@

The candidate genes were validated via the Gene Expression Profiling Interactive Analysis tool (GEPIA1/2 database using TCGA datasets, developed by Zefang Tang and his colleagues\cite{Tang2017a}\cite{Tang2019}

@@

Finally, the gene expression profile from the GEO and HNSCC cell lines and tissues were used to verify these biomarkers.\\

@@

%The ten genes, named DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, and DKC1, are significantly associated with the poor prognosis of overall survival (OS). 
%The other ten genes are over-expressed in the better survival group (with hazard ratio $<=0.6$), named as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (previous name as KIAA1683), and NPB.
%em dash ---
These CAMK2N1, CALML5, FCGBP and the other 17 genes (DKK1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1, ZNF557, ZNF266, MYO1H, LOC148709, EVPLL, PNMA5, IQCN---previous name as KIAA1683, and NPB) have significant FDR-adjusted \textit{P} values ($< 0.0003$) in Kaplan-Meier estimate, and greater hazard ratio (HR) ($> 1.8$ or $< 0.6$) in Cox's model (please see Figure \ref{fig:hazards3}; $log_{10}(0.0003) = -3.5$).
%These genes have also passed a stringent criteria (Bonferroni \textit{P} value correction less than 0.05).
The plot reveals that the top 20 genes (Bonferroni-adjusted \textit{P} $< 0.05$) are located on the peaks. At the same time, Cox's HR separates them on the two-side with significant prognostic impact.
%During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
%in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.

@@

%****************%
\subsubsection{Holistic Cancer Care}

@@

@@

% reviewer comment; Answer 2-4 (Cox regression)
\begin{MyColorPar}{red}
%The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival analysis data. It allows analyzing survival for one or more predictor variables and to provides the effect size (coefficients) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). Briefly, the hazard function can be interpreted as the risk of a specific event (e.g., death) at time t. It can be estimated as follow:

@@

$h(t) = h_0(t) \times exp(\beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n)$\\[0.3cm]
% natural exponential function => y=e^x
\end{flushleft}
where,\\
\begin{outline} % for highlight
\1 t represents the survival time;
\1 $h(t)$ is the hazard function determined by a set of n covariates ($X_1...X_n$, e.g., clinicopathological features: such as age, gender, gene expression, cancer stage, smoking, alcohol, or even spiritual, emotional, and social status);
\1 The coefficients ($\beta_1...\beta_n$) measure the impact (i.e., the effect size) of covariates;
\1 The term $h_0$ is called the baseline hazard. It corresponds to the hazard value if all the $X_i$ are equal to zero (the quantity exp(0) equals 1). The "t" in $h(t)$ indicates the hazard may vary over time.
\end{outline}

@@

\begin{MyColorPar}{red}
GSE65858 is a dataset for candidate selection in our workflow.
The Gene Expression Omnibus (GEO) database\cite{Chung2006},, founded by \acrfull{ncbi}, is a public repository supporting MIAME-compliant data submissions of microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007} was used to
download RNA-Seq dataset (in SOFT or MINiML format) of a HNSCC cohort, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?\\acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event, RFS event, and survival time.
There are 270 HNSCC participants involved in this cohort.
%The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30,330.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the platform GPL10558 (Illumina HumanHT-12 v4.0 Expression BeadChip), which targets on more than 30,330 annotated genes (47,000 probes, derived from the \acrshort{ncbi} Reference Sequence, release 38 on 7 November, 2009). % (RefSeq) 
GSE65858 is a dataset for helping of candidate selection in our workflow.
We have also performed Kaplan-Meier and Cox survival analyses under the cutoff of each gene expression at their median value.
\end{MyColorPar} % red

>>>>

\usepackage{xr-hyper}
\usepackage{hyperref}
%\newcommand{\R}[1]{\label{#1}\linelabel{#1}} % for \label page and line
\newcommand{\R}[1]{\linelabel{#1}} % for \label page and line
%\newcommand{\lr}[1]{page~\pageref{#1} (line~\lineref{#1})} % for \ref page and line
% for xr in your preamble
% https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
%\makeatletter
%\newcommand*{\addFileDependency}[1]{% argument=file name and extension
%  \typeout{(#1)}
%  \@addtofilelist{#1}
%  \IfFileExists{#1}{}{\typeout{No file #1.}}
%}
%\makeatother

%\newcommand*{\myexternaldocument}[1]{%
%    \externaldocument{#1}%
%    \addFileDependency{#1.tex}%
%    \addFileDependency{#1.aux}%
%}
% my external document I would like to reference the labels of. 
%\myexternaldocument{Supplementary_figures} %.tex .aux

@@

%\section{abstract}
%%%%%%%%%%
% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{The survival analysis of the Cancer Genome Atlas (TCGA) dataset is a well-known method to discover the gene expression-based prognostic biomarkers of head and neck squamous cell carcinoma (HNSCC). It is necessary to determine a cutoff point by the patients' dichotomization for the continuous gene expression in survival analysis. There is some optimization software for cutoff determination. However, the software's predetermined cutoffs are usually set at the median or quantiles of RNA sequencing value to perform the survival analyses. There are few clinicopathological features available on their pre-processed data sets.
We applied a in-house workflow, which includes data retrieving and pre-processing, feature selection, sliding-window cutoff selection, Kaplan-Meier survival analysis, and Cox proportional hazard modeling, for biomarker discovery.
Using this workflow on the TCGA HNSCC cohort, we scanned human protein-coding genes. After adjustment with confounders, we found that the clinical tumor stage and the surgical margin involvement are independent risk factors in patient survival. According to the resulting tables with Bonferroni adjusted P-value under the optimal cutoff and the hazard ratio, three biomarker candidates, CAMK2N1, CALML5, and FCGBP, are significantly associated with the patients' overall survival. We validated this discovery by using the other independent HNSCC dataset (GSE65858).
Thus, we suggest the transcriptomic analysis could help for biomarker discovery.}

@@

They used the Database for Annotation, Visualization, and Integrated Discovery (DAVID, available at https://david.ncifcrf.gov/) to obtain information for Gene Ontology (GO), including biological processes, the cellular component, and molecular function. 
The Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway analysis was also used to annotate their potential function.

@@

The candidate genes were validated by using the Gene Expression Profiling Interactive Analysis tool (GEPIA1/2 database using TCGA datasets, developed by Zefang Tang and his colleagues\cite{Tang2017a}\cite{Tang2019}

@@

Finally, these biomarkers were verified by using the gene expression profile from the GEO, HNSCC cell lines and tissues.\\

@@

%The ten genes, named DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, and DKC1, are significantly associated with the poor prognosis of overall survival (OS). 
%The other ten genes are over-expressed in the better survival group (with hazard ratio $<=0.6$), named as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (previous name as KIAA1683), and NPB.
%em dash ---
These CAMK2N1, CALML5, FCGBP and the other 17 genes (DKK1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1, ZNF557, ZNF266, IL19, MYO1H, EVPLL, PNMA5, IQCN, and NPB) have significant FDR-adjusted \textit{P} values ($< 0.0003$) in Kaplan-Meier estimate, and greater hazard ratio (HR) ($> 1.8$ or $< 0.6$) in Cox's model (please see Figure \ref{fig:hazards3}; $log_{10}(0.0003) = -3.5$).
%These genes have also passed a stringent criteria (Bonferroni \textit{P} value correction less than 0.05).
The plot reveals that the top 20 genes (Bonferroni-adjusted \textit{P} $< 0.05$) are located on the peaks. At the same time, Cox's HR separates them on the two-side with significant prognostic impact.
%During our exploration phase with TCGA cohort, the 20 genes, listed
%six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are bad guy;
%four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are good guy.
%}
%in Figure \ref{fig:hazards3}, have FDR-adjusted \textit{P} values ($<0.001$) in Kaplan-Meier estimate, and survival impact on HR ($> 1.8$ or $< 0.6$) in Cox's model.

@@

% in coverletter
% line \ref{line_fr:answer1-2} ~ \ref{line_to:answer1-2} of page %\pageref{page_fr:answer1-2}, \pageref{page_to:answer1-2}
% in main.tex
% %\label{page_fr:answer1-2}, \label{page_to:answer1-2}
% \linelabel{line_fr:answer1-2} , \linelabel{line_to:answer1-2}
%****************%
\subsubsection{Holistic Cancer Care} 
\R{from:answer1-2}
%\linelabel{line_fr:answer1-2}

@@

\R{to:answer1-2}
%\label{page_to:answer1-2} % page number
%\linelabel{line_to:answer1-2} % line number
% \label{line_to:answer1-2} % section number

@@

% reviewer comment; Answer 2-4 (Cox regression)
\begin{MyColorPar}{red}
%The Cox proportional-hazards regression model\cite{Cox1972}\cite{Andersen1982} is commonly used for modeling survival analysis data. It allows analyzing survival for one or more predictor variables and to provides the effect size (coefficients) for them\cite{Bradburn2003b}. % Classification trees and artificial neural networks in survival analysis
The Cox model is also accounting for confounding factors\cite{Magen2019}.
It is expressed by the hazard function denoted by h(t). 
The hazard function represents the risk of a specific event (e.g., death) at time t. It can be estimated as follow:

@@

$h(t) = h_0(t) \times exp(\beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + ... + \beta_n X_n)$\\[0.3cm]
% natural exponential function => y=e^x
\end{flushleft}
where,\\
\begin{outline} % for highlight
\1 t represents the survival time;
\1 $h(t)$ is the hazard function determined by a set of n covariates ($X_1...X_n$, e.g., clinicopathological features, including age, gender, gene expression, cancer stage, smoking, alcohol, spiritual, emotional, and social status;
\1 The coefficients ($\beta_1...\beta_n$) measure the impact---the effect size of covariates;
\1 The term $h_0$ is called the baseline hazard. It corresponds to the hazard value if all the $X_i$ are equal to zero. The "t" in $h(t)$ indicates the hazard may vary over time.
\end{outline}

@@

\begin{MyColorPar}{red}
GSE65858 is a dataset for candidate selection in our workflow.
The Gene Expression Omnibus (GEO) database\cite{Chung2006},, founded by \acrfull{ncbi}, is a public repository supporting MIAME-compliant data,  including microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007} was used to
download RNA-Seq dataset (in SOFT or MINiML format) of a HNSCC cohort, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?\\acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event, RFS event, and survival time.
There are 270 HNSCC participants involved in this cohort.
%The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30,330.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the platform GPL10558 (Illumina HumanHT-12 v4.0 Expression BeadChip), which targets on more than 30,330 annotated genes (47,000 probes, derived from the \acrshort{ncbi} Reference Sequence, release 38 on 7 November, 2009). % (RefSeq) 
GSE65858 is a dataset for helping of candidate selection in our workflow.
We have also performed Kaplan-Meier and Cox survival analyses under the cutoff of each gene expression at their median value.
\end{MyColorPar} % red
